---
title: "Manuscript Figures"
format: html
editor: visual
---

## This is code for generating all figures in the MS.

**NOTE**: **Please restart R before running any code in this file and do not click RUN for this document at the top of the script**. There is a bug with ecodist that will be triggered if you do not restart R or do not run cell by cell as a function will be overwritten by the proxy package: <https://github.com/phiala/ecodist>

```{r load libraries}
library(bipartite)
library(broom)
library(car)
library(cheddar)
library(dplyr)
library(forcats)
library(ggraph)
library(ggrepel)
library(ggmap)
library(ggplot2)
library(ggspatial)
library(igraph)
library(lmtest)
library(MASS)
library(Matrix)
library(metanetwork)
library(patchwork)
library(phytools)
library(rlang)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(spdep)
library(spatialreg)
library(tidygraph)
library(tidyr)
library(tidyverse)
library(tseries)
library(units)
library(vegan)
library(ecodist)


# This confirms if the ecodist package is working
dim(dist(matrix(1:12, ncol = 3)))   # should return NULL, if not then proxy is overwriting ecodist
```

**Due to issues with ecodist, this cell must be run first.**

```{r AppendixS4 Multi Mantel}


# Align the FGW matrix as needed for these 127 sites
fgw_matrix_function <- function(useobject) {
  
  # Get names of sites to use as rownames in matrix to subset FGW 
  sitenames <- useobject$Code
  rownames(useobject) <- sitenames
  
  
  # Bring in site data to use later for rownames
  SiteTable_raw <- read.table(file="African_Mammal_FoodWebs_Locations.csv", header=TRUE, sep=",")
  
  # Bring in fused GW distance matrix
  FW_approach_raw <- read.csv("pairwise_fgw.csv", sep=",")
  # Remove rownames row
  FW_approach_raw <- FW_approach_raw[,-1]
  
  # Match order of locations and FGW csv to correctly add codes for subsetting by region later
  SiteTable <- SiteTable_raw[match(colnames(FW_approach_raw), SiteTable_raw$Community), ]
  
  
  # Remove colnames and rownames for proper formatting later
  colnames(FW_approach_raw) <- NULL
  FW_approach <- data.matrix(FW_approach_raw)
  
  # Changing to site code names to match later to env variables
  rownames(FW_approach) <- SiteTable$Code
  colnames(FW_approach) <- SiteTable$Code
  # Create overall distance matrix in the proper format for the function
  comm <- FW_approach[match(rownames(useobject),rownames(FW_approach)),match(rownames(useobject),rownames(FW_approach))]
  dim(comm)
  comm <- comm[,colSums(comm, na.rm=TRUE) > 0]
  dim(comm)
  overall_matrix_raw <- as.matrix(comm)  
  
  # Now we will take the log of these distances, see appendix for info for how this accounts for
  # the impact of # of species on the FGW.
  overall_matrix <- log(overall_matrix_raw)
  
  # This will replace negative values that results from very small FGW caused by being the same web
  # This happens in only 1 site
  overall_matrix[overall_matrix < 0] <- 0
  
  # Convert to similarity
  #overall_matrix <- overall_matrix*-1
  
  return(overall_matrix)
}


multi_mantel_regions <- function(region_name) {
  
  # Bring in all predictors
  all_env_vars_raw <- read.table(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")

  # Only select the sites to be included in the analysis
  all_env_vars <- subset(all_env_vars_raw, Remove==0)
  
  if (region_name=="All") {
    sites_to_use <- all_env_vars 
  } else {
    sites_to_use <- subset(all_env_vars, Region==region_name)
  }


  # Create distance matrix of predictors
  prod_dist <- dist(scale(sites_to_use$EVI_mean),method = "euclidean",diag = TRUE, upper = TRUE) 

  #Subtract agg from 100 to get a value for fragmentation
    frag_dist <- dist(scale(100-sites_to_use$bin.AG.class),method = "euclidean",diag = TRUE, upper = TRUE)  # Euclidean distance

  # Make coordinates into spatial objects
  coords <- data_frame(Long=sites_to_use$Long, Lat=sites_to_use$Lat)

  # Change to meters
  sites <- st_as_sf(coords, coords = c("Long", "Lat"), crs = 4326)
  dist_m <- st_distance(sites)   # units = meters
  geo_dist_unit <- units::set_units(dist_m, "km") # change unit to kilometers

  # Create distance matrix
  geo_dist <- drop_units(geo_dist_unit)
  

  # Subet fGW for these sites 
  fgw <- fgw_matrix_function(sites_to_use)
  
  
  ## RUN MODELS ##
  # Initialize list to save results 
  results <- vector(mode = "list", length = 8)
  names(results) = c('EVI','Frag','Geo', 'R-Squared MRM','R-Squared Hand-Calc', 'lm R-Squared', 'Veg Plot', 'Frag Plot')
  # Set seed for reproducibilty
  set.seed(123)
  # EVI
 evi <-  ecodist::mantel(as.dist(fgw) ~ as.dist(prod_dist) + as.dist(frag_dist) + as.dist(geo_dist),nperm = 10000)
 
 # Save results to list
 results[[1]] <- evi

  # Frag
  frag <- ecodist::mantel(as.dist(fgw) ~ as.dist(frag_dist) + as.dist(prod_dist) + as.dist(scale(geo_dist)),nperm = 10000)
  
  # Save results to list
 results[[2]] <- frag

  # Geo
  geo <- ecodist::mantel(as.dist(fgw) ~ as.dist(geo_dist) + as.dist(prod_dist) + as.dist(frag_dist),nperm = 10000)
  
  # Save results to list
 results[[3]] <- geo
  
  # Now let's try to calculate R-squared
  r_sq <- MRM(formula = as.dist(fgw) ~ as.dist(prod_dist) + as.dist(frag_dist) + as.dist(geo_dist), nperm = 1000,method = "linear", mrank = FALSE)
  
  
  # Save results to list
 results[[4]] <- r_sq
  
  # Calculate R-Squared a different way
  r_y1 <- ecodist::mantel(as.dist(fgw) ~ as.dist(prod_dist))[[1]] # Mantel r between Y and X1
  r_y2 <- ecodist::mantel(as.dist(fgw) ~ as.dist(frag_dist))[[1]]
  r_y3 <- ecodist::mantel(as.dist(fgw) ~ as.dist(geo_dist))[[1]]

  r_12 <- ecodist::mantel(as.dist(prod_dist) ~ as.dist(frag_dist))[[1]]
  r_13 <- ecodist::mantel(as.dist(prod_dist) ~ as.dist(geo_dist))[[1]]
  r_23 <- ecodist::mantel(as.dist(frag_dist) ~ as.dist(geo_dist))[[1]]

  r_yX <- c(r_y1, r_y2, r_y3)
  R_XX <- matrix(c(1,   r_12, r_13,
                 r_12, 1,   r_23,
                 r_13, r_23, 1),
               nrow=3, byrow=TRUE)

  # invert (use ginv if singular)
  Rinv <- tryCatch(solve(R_XX), error = function(e) ginv(R_XX))

  R2_from_mantel <- as.numeric(t(r_yX) %*% Rinv %*% r_yX)
  # Save results to list
 results[[5]] <- R2_from_mantel
  
   
 
  # Finally, visualize
  # Vectorize distance matrices and multiply by -1 to create similarity
  y_vec <- as.vector(as.dist(fgw)*-1)
  geo_vec <- as.vector(as.dist(geo_dist)*-1)
  veg_vec <- as.vector(as.dist(prod_dist)*-1)
  frag_vec <- as.vector(as.dist(frag_dist)*-1)
  
  # Check R Squared in stsnard lm
   lm_model <- summary(lm(y_vec ~ geo_vec + veg_vec + frag_vec))
   results[[6]] <- lm_model 

  # Regress out controls to plot veg
  res_fgw_veg <- residuals(lm(y_vec ~ frag_vec + geo_vec))
  res_veg   <- residuals(lm(veg_vec ~ frag_vec + geo_vec))
  # Plot
  veg_resid_plot <- ggplot(data.frame(res_veg, res_fgw_veg), aes(x = res_veg, y = res_fgw_veg)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method="lm", color="blue", se=FALSE) +
    labs(x = "Veg residuals", y = "FGW residuals") +
    theme_minimal() 
  
  
  # Regress out controls to plot frag
  res_fgw_frag <- residuals(lm(y_vec ~ veg_vec + geo_vec))
  res_frag   <- residuals(lm(frag_vec ~ veg_vec + geo_vec))
  # Plot
  frag_resid_plot <- ggplot(data.frame(res_frag, res_fgw_frag), aes(x = res_frag, y = res_fgw_frag)) +
    geom_point(alpha = 0.4) +
    geom_smooth(method="lm", color="blue", se=FALSE) +
    labs(x = "Frag residuals", y = "FGW residuals") +
    theme_minimal() 
  
  
  # Save results to list
 results[[7]] <- veg_resid_plot
 results[[8]] <- frag_resid_plot
 
  
  return(results)

  
}

multi_mantel_regions("All")
multi_mantel_regions("Congolian Forests")
multi_mantel_regions("Lake Turkana")
multi_mantel_regions("Southern Drylands")
multi_mantel_regions("Subequatorial Savannas & Woodlands")
multi_mantel_regions("Sudanian Savanna")
multi_mantel_regions("Victoria Basin & Albertine Rift Forests")






```

## Main text Figures

```{r Figure2 Metaweb}
# Bring in metaweb csv
metaweb_raw <- read_csv("Africa_metaweb_above500.csv")
# Change rownames to prey species column
metaweb_names <- tibble::column_to_rownames(metaweb_raw, var = "Prey_Species") 
# Change to matrix
metaweb <- as.matrix(metaweb_names)

# Bring in presence/absence data
species_present_all <- read_csv("above_500_PA_data.csv")

# Remove species column
species_present_raw <- species_present_all[,-1]
# Switch columns and rows of species present to get data in correct format for metaweb
species_present <- (t(species_present_raw))
colnames(species_present) <- species_present_all$Species


# Creating S3 metanetwork object
meta_Africa <- build_metanet(
  metaweb,
  abTable=species_present
)

# Check an individual site
#plot(meta_Africa$ABER)
#plot(meta_Africa$KFR)


# Let's make a better lookinh metaweb with mbetter labeling
#https://jeremydfoote.com/Communication-and-Social-Networks/resources/ggraph_walkthrough.html
metaweb_nicer <- read_csv("Africa_metaweb_above500.csv")
# Change rownames to prey species column
metaweb_nicer <- column_to_rownames(metaweb_nicer, var = "Prey_Species") %>% as_tibble()

# Change to matrix
metaweb_nicer_matrix <- as.matrix(metaweb_nicer)

# Bring in node information
node_info <- read.csv("Node_Data_500g.csv")

# Creating graph object from adjacency matrix and turning it into a tidy graph
metaweb_nicer_adj <- graph_from_adjacency_matrix(metaweb_nicer_matrix,mode="directed",diag=FALSE) 

metaweb_nicer_graph <- as_tbl_graph(metaweb_nicer_adj)

# Adding node information
metaweb_nicer_graph  <- metaweb_nicer_graph  %>% 
  activate(nodes) %>% 
  mutate(mass=node_info$Mass.g, order=node_info$Order)

# Check if it worked
metaweb_nicer_graph

# Graph it
meta_tg <- metaweb_nicer_graph %>% 
  ggraph(layout = "fr") +
  geom_edge_arc(colour= "black",
                lineend = "round",
                strength = .1,
                alpha = .1) +
  geom_node_point(aes(color=order, size=log(mass/1000))) +
  geom_node_text(aes(label = name), 
                 repel = TRUE, 
                 point.padding = unit(0.1, "lines"),
                 size=3,
                 colour="black") +
  scale_size_continuous(name = "Log(Body Mass in kg)") +
  theme_graph(background = "white") +
  scale_color_manual(name = "Order",values = c("red", "navyblue", "green","purple","black",
                                "brown","pink","orange","lightblue",
                                "gray","darkgreen","yellow")) +
  guides(edge_width = FALSE,
         edge_alpha = FALSE)

meta_tg




# Remove species labels 
meta_tg_unlabeled <- metaweb_nicer_graph %>% 
  ggraph(layout = "fr") +
  geom_edge_arc(colour= "black",
                lineend = "round",
                strength = .1,
                alpha = .1) +
  geom_node_point(aes(color=order, size=log(mass/1000))) +
  scale_size_continuous(name = "Log(Body Mass in kg)") +
  theme_graph(background = "white") +
  scale_color_manual(name = "Order",values = c("red", "navyblue", "green","purple","black",
                                "brown","pink","orange","lightblue",
                                "gray","darkgreen","yellow")) +
  guides(edge_width = FALSE,
         edge_alpha = FALSE) + 
  theme(legend.text = element_text(size = 12),  # Adjust the size for legend text
        legend.title = element_text(size = 14)) # Adjust the size for legend title

meta_tg_unlabeled
```

```{r Figure3 Map of Sites}
# Load Africa map data
africa <- ne_countries(scale = "medium", returnclass = "sf", continent = "Africa")
africa <- africa[africa$admin != "Madagascar", ]  # Filter out Madagascar by name

# Bring in data 
map_info_raw <- read.csv(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")
# Remove excluded sites
map_info_sites_to_use <- subset(map_info_raw, Remove==0)

# Create data frame of points
data_points <- data.frame(
  lon = map_info_sites_to_use$Long,  # Pull longitude coordinates from dataset
  lat = map_info_sites_to_use$Lat,      # Pull latitude coordinates from dataset
  Region = map_info_sites_to_use$Region # regions from dataset
)


# Plot the map with scale bar and north arrow
ggplot(data = africa) +
  geom_sf(fill = "gray90", color = NA) +  # Plot Africa map
  geom_point(data = data_points, aes(x = lon, y = lat, shape = Region), size = 2) +
  scale_shape_manual(values = c(3, 8, 5, 16, 15, 17)) +  # Define shapes for Regions
  labs(
    title = "Map of Study Sites",
    x = "Longitude",
    y = "Latitude",
    color = "Region"
  ) +
  # Add scale bar
  #annotation_scale(location = "bl", width_hint = 0.2) +
  # Add north arrow
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.5, "cm"), pad_y = unit(0.75, "cm"), # Adjust padding to move north arrow up
    style = north_arrow_fancy_orienteering
  ) + 
    # Remove gridlines
  theme_minimal() +
  theme(
    panel.grid = element_blank(),   # Remove gridlines
    panel.background = element_rect(fill = "white", color = NA), # White background
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.line = element_line(color = "black", size = 0.5)  # Add axis lines
  )
```

```{r Figure5 All Site Added Variable Plot}

# Align the FGW matrix as needed for these 127 sites
fgw_matrix_function <- function(useobject) {
  
  # Get names of sites to use as rownames in matrix to subset FGW 
  sitenames <- useobject$Code
  rownames(useobject) <- sitenames
    # Bring in site data to use later for rownames
  SiteTable_raw <- read.table(file="African_Mammal_FoodWebs_Locations.csv", header=TRUE, sep=",")
  
  # Bring in fused GW distance matrix
  FW_approach_raw <- read.csv("pairwise_fgw.csv", sep=",")
  # Remove rownames row
  FW_approach_raw <- FW_approach_raw[,-1]
  
  # Match order of locations and FGW csv to correctly add codes for subsetting by region later
  SiteTable <- SiteTable_raw[match(colnames(FW_approach_raw), SiteTable_raw$Community), ]
  
  
  # Remove colnames and rownames for proper formatting later
  colnames(FW_approach_raw) <- NULL
  FW_approach <- data.matrix(FW_approach_raw)
  
  # Changing to site code names to match later to env variables
  rownames(FW_approach) <- SiteTable$Code
  colnames(FW_approach) <- SiteTable$Code
  
  
  
  # Create overall distance matrix in the proper format for the function
  comm <- FW_approach[match(rownames(useobject),rownames(FW_approach)),match(rownames(useobject),rownames(FW_approach))]
  dim(comm)
  comm <- comm[,colSums(comm, na.rm=TRUE) > 0]
  dim(comm)
  overall_matrix_raw <- as.matrix(comm)  
  
  # Now we will take the log of these distances, see appendix for info for how this accounts for
  # the impact of # of species on the FGW.
  overall_matrix <- log(overall_matrix_raw)
  
  # This will replace negative values that results from very small FGW caused by being the same web
  # This happens in only 1 site
  overall_matrix[overall_matrix < 0] <- 0
  
  # Convert to similarity
  #overall_matrix <- overall_matrix*-1
  
  return(overall_matrix)
}



all_env_vars_raw <- read.table(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")

# Only select the sites to be included in the analysis
all_env_vars <- subset(all_env_vars_raw, Remove==0)


# Create distance matrix of predictors
prod_dist_all <- dist(scale(all_env_vars$EVI_mean),method = "euclidean",diag = FALSE, upper = FALSE) 

frag_dist_all <- dist(scale(100-all_env_vars$bin.AG.class),method = "euclidean",diag = FALSE, upper = FALSE)  # Euclidean distance

# Make coordinates into spatial objects
coords <- data_frame(Long=all_env_vars$Long, Lat=all_env_vars$Lat)

# Change to meters
sites <- st_as_sf(coords, coords = c("Long", "Lat"), crs = 4326)
dist_m <- st_distance(sites)   # units = meters
geo_dist_unit <- units::set_units(dist_m, "km") # change unit to kilometers

# Create distance matrix
geo_dist <- drop_units(geo_dist_unit)

# Subet fGW for these sites 
fgw_all <- fgw_matrix_function(all_env_vars)



# Okay now let's plot this in a nice plot
# Finally, visualize
# Vectorize distance matrices and multiply by -1 to create similarity
fgw_vec <- as.vector(as.dist(fgw_all)*-1)
geo_vec <- as.vector(as.dist(scale(geo_dist))*-1)
veg_vec <- as.vector(as.dist(prod_dist_all)*-1)
frag_vec <- as.vector(as.dist(frag_dist_all)*-1)

# Get the names of the sites 
site_names <- colnames(fgw_all)
# Create a pairwise list
pairwise_list <- combn(site_names, 2, simplify = FALSE)
pairwise_df <- as.data.frame(do.call(rbind, pairwise_list))
colnames(pairwise_df) <- c("Site1", "Site2")

# Combine the two columns into one with names separated by "_"
pairwise_df$Combined <- paste(pairwise_df$Site1, pairwise_df$Site2, sep = "_")

# Add vectors to dataframe
all_model <- data.frame(pairwise_df$Combined,veg_vec, frag_vec, geo_vec, fgw_vec)

# Bring in region data for labeling later
region_info_raw <- read.csv(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")
# Remove excluded sites
region_info_to_use <- subset(region_info_raw, Remove==0)

# Grab columns we need, Code and Region
region_info <- region_info_to_use[,c(1,2)]

# Split pair into two site columns
all_model <- all_model %>%
  separate(pairwise_df.Combined, into = c("site1", "site2"), sep = "_")

# Add region info for each site
all_model <- all_model %>%
  left_join(region_info, by = c("site1" = "Code")) %>%
  rename(region1 = Region) %>%
  left_join(region_info, by = c("site2" = "Code")) %>%
  rename(region2 = Region)

# Create SAME/DIFFERENT column
all_model <- all_model %>%
  mutate(region_match = ifelse(region1 == region2, "SAME", "DIFFERENT"))

all_model <- all_model %>%
  mutate(shared_region = ifelse(region1 == region2, region1, "No region"))

# Make column into vector of colors
cols <- ifelse(all_model$region_match == "SAME", "purple", "black")

# Make column into vector of shapes
shapes <- ifelse(all_model$region_match == "SAME", 16, 1)







# Set up plotting area: 1 row, 2 columns
par(mfrow = c(1, 2))

# First plot: EVI
model <- lm(fgw_vec ~ veg_vec + frag_vec + geo_vec, data = all_model)
avPlot(model, "veg_vec",
       main = "",
       xlab = "Plant Productivity Similarity", ylab = "Food Web Similarity",
       ylim = c(-3, 3),
       id = FALSE,
       grid = FALSE,
       col=cols,
       pch = shapes) 

legend("bottomright", 
       legend = c("Same Region", "Different Region"),
       col = c("purple", "black"),
       pch = c(16, 1),
       bty = "n")

# First plot label
text(x = par("usr")[1] + 0.05 * diff(par("usr")[1:2]),
     y = par("usr")[4] - 0.05 * diff(par("usr")[3:4]),
     labels = "a.", cex = 1.2, font = 2)


# Second plot: Fragmentation
model <- lm(fgw_vec ~ veg_vec + frag_vec + geo_vec, data = all_model)
avPlot(model, "frag_vec",
       main = "",
       xlab = "Fragmentation Similarity", ylab = "Food Web Similarity",
       ylim = c(-3, 3),
       col.lines = NA,
       id = FALSE,
       grid = FALSE,
       col=cols,
       pch = shapes)
abline(lm(residuals(model) ~ residuals(lm(frag_vec ~ veg_vec + geo_vec, data = all_model))),
       lty = 2, lwd = 2, col="blue")  

legend("bottomright", 
       legend = c("Same Region", "Different Region"),
       col = c("purple", "black"),
       pch = c(16, 1),
       bty = "n")


# Second plot label
text(x = par("usr")[1] + 0.05 * diff(par("usr")[1:2]),
     y = par("usr")[4] - 0.05 * diff(par("usr")[3:4]),
     labels = "b.", cex = 1.2, font = 2)


  

  
```

```{r Figure6 Partial Mantel and Predictors Visualization}

# Partial Mantel Region Results
# Bring in partial mantel data 
pm_results_barplot_raw <- read.csv(file="Partial_Mantel_Barplot.csv", header=TRUE, sep=",")

# Remove the all_sites results
pm_results_barplot <- subset(pm_results_barplot_raw, Region!="All")

# Placing them in order of increasing productivity
desired_order <- c("DRY","LAKET","SDSVN","SVN","VBAR","CONGO")
pm_results_barplot <- pm_results_barplot %>%
  mutate(Region = factor(Region, levels = desired_order))


all_var_plot <- ggplot(pm_results_barplot, aes(fill=Predictor, y=r, x=Region)) + 
    geom_bar(position="dodge", stat="identity") + 
  theme(panel.grid = element_blank(), 
        axis.title.x = element_text(vjust = 0, size = 15),
        axis.title.y = element_text(vjust = 0, size = 15),
        axis.text = element_text(size = 12),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  scale_fill_manual(values=c( "chocolate4","green4")) +
  ylab("Partial Mantel correlation coefficient (r)") +
  scale_y_continuous(limits = c(-0.2, 0.5)) + 
  theme(
    legend.position = c(0.05, 0.95),  # x = left, y = top
    legend.justification = c("left", "top"),  
    legend.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  ) + annotate(
    "text", x = -Inf, y = Inf, label = "(b)", 
    size = 6, fontface = "bold", hjust = -0.1, vjust = 1
  )


# Variation in Predictors
predictor_info <- read.csv(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")
# Creating new dataframe to sort for EVI order in violin plot
evi_regions <- subset(predictor_info, Remove==0)

# Labels in order of ascending EVI
evi_labels=c("DRY","LAKET","SDSVN","SVN","VBAR","CONGO")

# Plot
p_evi <- evi_regions %>% 
  mutate(Region = fct_reorder(Region, EVI_mean, .fun = mean)) %>% 
  ggplot(aes(Region, EVI_mean)) + 
  geom_violin(width = 1) + labs(
    x = "Region",
    y = "Primary Productivity (EVI)",
    title = "") + 
  theme(panel.grid = element_blank(), 
        axis.title.x = element_text(vjust = 0, size = 15),
        axis.title.y = element_text(vjust = 0, size = 15),
        axis.text = element_text(size = 12),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  geom_boxplot(width=0.1) +  
  scale_x_discrete(labels=evi_labels) + 
  annotate(
    "text", x = -Inf, y = Inf, label = "(a)", 
    size = 6, fontface = "bold", hjust = -0.1, vjust = 1
  )

# EVI Values Summary
evi_regions  %>% group_by(Region) %>%
summarize(mean(EVI_mean))



# Plot
p_evi + all_var_plot
```

## Appendix Figures

```{r AppendixS1 Metaweb Link Information}
# We added more links after combining data. The following numbers denote the same type of link (Documented (1 & 4), Genus (2 & 5), and Mass based (3 & 6). 1-3 are from our original data collection, and 4-6 were added later. We combined these sums in the appendix summary for each type of link. 

# Bring in links
new_links_raw <- read_csv("Africa_metaweb_above500.csv")
# Change rownames to prey species column
new_links_raw <- column_to_rownames(new_links_raw, var = "Prey_Species") %>% as_tibble()
# Change to matrix
new_links <- as.matrix(new_links_raw)
# Bring in node information
node_info <- read.csv("Node_Data_500g.csv")
# Creating graph object from adjacency matrix and turning it into a tidy graph
new_links_adj <- graph_from_adjacency_matrix(new_links,mode="directed",diag=FALSE,weighted=TRUE) 
new_links_adj <- as_tbl_graph(new_links_adj)
# Adding node information
meta_new_links  <- new_links_adj  %>% 
  activate(nodes) %>% 
  mutate(mass=node_info$Mass.g, order=node_info$Order)


# Between Original Food Web and Evan Fricke, we added 320 new carnivore links

# This gives us all the info about the web in a dataframe
web_info_df = as_long_data_frame(meta_new_links)

# Lets see how many of each weight there is, aka how many of each type of link.
web_info_df  %>% group_by(weight) %>%
summarize(n())
# These numbers are summed up in appendix S2, table A.



# Now let's add "from_order" because this will tell us how many link types from prey
# Prey Tallies
web_info_df %>%
  mutate(weight_group = case_when(
    weight %in% c(1, 4) ~ "Documented",
    weight %in% c(2, 5) ~ "Genus",
    weight %in% c(3, 6) ~ "Mass"
  )) %>%
  group_by(weight_group, from_order) %>%
  summarize(count = n(), .groups = "drop") %>%
  complete(weight_group, from_order, fill = list(count = 0)) %>%
  pivot_wider(
    names_from = weight_group,
    values_from = count
  )

# Now let's add "to_order" because this will tell us how many link types from predator
# Predator Tallies
web_info_df %>%
  mutate(weight_group = case_when(
    weight %in% c(1, 4) ~ "Documented",
    weight %in% c(2, 5) ~ "Genus",
    weight %in% c(3, 6) ~ "Mass"
  )) %>%
  group_by(weight_group, to_order) %>%
  summarize(count = n(), .groups = "drop") %>%
  complete(weight_group, to_order, fill = list(count = 0)) %>%
  pivot_wider(
    names_from = weight_group,
    values_from = count
  )


```

```{r AppendixS2 Justification for logging fGW}
# Get the distance matrices for number of nodes and fused GW distance to plot later
FW_approach_raw <- read.csv("pairwise_fgw.csv", sep=",")

# Bring in site data to use later for rownames
SiteTable <- read.table(file="African_Mammal_FoodWebs_Locations.csv", header=TRUE, sep=",")

# Remove colnames and rownames 
FW_approach_raw <- FW_approach_raw[,-1]
colnames(FW_approach_raw) <- NULL
FW_approach <- data.matrix(FW_approach_raw)

# Changing to site code names
rownames(FW_approach) <- SiteTable$Code
colnames(FW_approach) <- SiteTable$Code

# Bring in env data
all_env_vars_raw <- read.table(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")

# Only select the sites included in the analysis
all_env_vars <- subset(all_env_vars_raw,Remove==0)

# Select just the predictor variables we want
env1 <- scale(all_env_vars$num_nodes)
env2 <- all_env_vars$num_nodes


# Combine the env variable that we are using 
env_vars <- data.frame(Code=all_env_vars$Code,env1,env2)
# Set column names to which variables are being used
colnames(env_vars) <- c("Code","Scaled_Num_Nodes","Num_Nodes")

# Remove site code 
env_sites <- subset(env_vars, select=-c(Code))

# Get names of sites to use as rownames in matrix 
sitenames <- env_vars$Code
rownames(env_sites) <- sitenames


# Create overall distance matrix matching site names
comm <- FW_approach[match(rownames(env_sites),rownames(FW_approach)),match(rownames(env_sites),rownames(FW_approach))]
dim(comm)
comm <- comm[,colSums(comm, na.rm=TRUE) > 0]
dim(comm)
overall_matrix <- as.matrix(comm)  


# Get number of nodes difference

# Using dist function to get difference
node_difference <- dist(env_sites$Num_Nodes,diag = FALSE, upper = FALSE)

# Change to vector
node_difference_vector <- as.vector(node_difference)

# FGW to match, need to remove the diagonal
fgw_vector_to_match <- as.vector(as.dist(overall_matrix))


# Plot
par(mfrow=c(1,2))
plot(node_difference_vector,fgw_vector_to_match, main="Difference in Number of Nodes vs. Fused GW",
     xlab="Difference in Number of Nodes", ylab="Fused GW Distance")
plot(node_difference_vector,log(fgw_vector_to_match), main="Difference in Number of Nodes vs. Logged Fused GW",ylim=c(5,15),
     xlab="Difference in Number of Nodes", ylab="Logged Fused GW Distance")


###### Now let's look at connectance ########


# Get the distance matrices for number of nodes and fused GW distance to plot later
FW_approach_raw <- read.csv("pairwise_fgw.csv", sep=",")

# Bring in site data to use later for rownames
SiteTable <- read.table(file="African_Mammal_FoodWebs_Locations.csv", header=TRUE, sep=",")

# Remove colnames and rownames 
FW_approach_raw <- FW_approach_raw[,-1]
colnames(FW_approach_raw) <- NULL
FW_approach <- data.matrix(FW_approach_raw)

# Changing to site code names
rownames(FW_approach) <- SiteTable$Code
colnames(FW_approach) <- SiteTable$Code

# Bring in data on which regions to keep
all_env_vars_raw_input <- read.table(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")
network_metrics_raw <- read.table(file="NetworkMetrics.csv", header=TRUE, sep=",")
#Drop remove column
network_metrics_raw <- network_metrics_raw[,-4]
  
all_env_vars_raw <- merge(all_env_vars_raw_input,network_metrics_raw, by="Code" )
# Only select the sites included in the analysis
all_env_vars <- subset(all_env_vars_raw,Remove==0)


# Select just the predictor variables we want
conect <- all_env_vars$Connectance


# Combine the two env variable that we are using 
env_vars <- data.frame(Code=all_env_vars$Code,conect)
# Set column names to which variables are being used
colnames(env_vars) <- c("Code","Connectance")

# Remove site code 
env_sites <- subset(env_vars, select=-c(Code))

# Get names of sites to use as rownames in matrix 
sitenames <- env_vars$Code
rownames(env_sites) <- sitenames


# Create overall distance matrix matching site names
comm <- FW_approach[match(rownames(env_sites),rownames(FW_approach)),match(rownames(env_sites),rownames(FW_approach))]
dim(comm)
comm <- comm[,colSums(comm, na.rm=TRUE) > 0]
dim(comm)
overall_matrix <- as.matrix(comm)  


# Using dist function to get difference
connectance_difference <- stats::dist(env_sites$Connectance,diag = FALSE, upper = FALSE,method = "euclidean")
# Change to vector to plot
connectance_vector <- as.vector(connectance_difference)


# FGW to match
fgw_vector_to_match <- as.vector(as.dist(overall_matrix))


# Plot
par(mfrow=c(1,2))
plot(connectance_vector,fgw_vector_to_match, main="Difference in Connectance vs. Fused GW",
     xlab="Difference in Connectance", ylab="Fused GW Distance")

plot(connectance_vector,log(fgw_vector_to_match), main="Difference in Connectance vs. Logged Fused GW",
     xlab="Difference in Connectance", ylab="Logged Fused GW Distance", ylim=c(5,15))



# Calculate correlations between fgw, species richness, and connectance 
cor(connectance_vector,fgw_vector_to_match)
cor(node_difference_vector,fgw_vector_to_match)


```

```{r AppendixS3}
# See GEE scripts for how to get raw data 
predictor_info <- read.csv(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")

predictor_info <- subset(predictor_info, Remove==0)

# Biogeographic Region Grouped
cluster1 <- subset(predictor_info,Region=="Lake Turkana")
cluster2 <- subset(predictor_info,Region=="Southern Drylands")
cluster3 <- subset(predictor_info,Region=="Victoria Basin & Albertine Rift Forests")
cluster4 <- subset(predictor_info,Region=="Sudanian Savanna")
cluster5 <- subset(predictor_info,Region=="Subequatorial Savannas & Woodlands")
cluster6 <- subset(predictor_info,Region=="Congolian Forests")

# Standard deviation within regions
evi_sd <- predictor_info %>%
  group_by(Region) %>%
  summarise(std_value = mean(EVI_std, na.rm = TRUE))

evi_mean <- predictor_info %>%
  group_by(Region) %>%
  summarise(mean_value = mean(EVI_mean, na.rm = TRUE))

# Coefficent of variance
(evi_sd$std_value/evi_mean$mean_value)*100
  
# Repeat for all sites
(mean(predictor_info$EVI_std, na.rm = TRUE)/mean(predictor_info$EVI_mean, na.rm = TRUE))*100


# Mean value within regions
mean(predictor_info$EVI_mean, na.rm = TRUE)

# Again now for fragmentation measures
# STandard deviation within regions
frag_sd <- predictor_info %>%
  group_by(Region) %>%
  summarise(std_value = mean(Frag_std, na.rm = TRUE))

frag_mean <- predictor_info %>%
  group_by(Region) %>%
  summarise(mean_value = mean(bin.AG.class, na.rm = TRUE))

# Coefficent of variance
(frag_sd$std_value/frag_mean$mean_value)*100
  
# Repeat for all sites
(mean(predictor_info$Frag_std, na.rm = TRUE)/mean(predictor_info$bin.AG.class, na.rm = TRUE))*100


# Mean value within regions

```

```{r AppendixS5 Fragmentation Plot}

# bring in data on variation in Predictors
predictor_info <- read.csv(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")

# Creating new dataframe to sort for Frag order in violin plot
frag_regions <- subset(predictor_info, Remove==0)

# Labels in order for Frag
frag_labels=c("DRY","LAKET","SDSVN","SVN","VBAR","CONGO")

p_frag <- frag_regions %>% 
  mutate(Region = fct_reorder(Region, EVI_mean, .fun = mean)) %>% 
  ggplot(aes(Region, 100-bin.AG.class)) + 
  geom_violin(width = 1) + labs(
    x = "Region",
    y = "Binary Fragmentation Index %",
    title = "") + 
  theme(panel.grid = element_blank(), 
        axis.title.x = element_text(vjust = 0, size = 15),
        axis.title.y = element_text(vjust = 0, size = 15),
        axis.text = element_text(size = 12),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black")) + 
  geom_boxplot(width=0.1) + 
  scale_x_discrete(labels=frag_labels) 

# Frag Values Summary
frag_regions  %>% group_by(Region) %>%
summarize(100-mean(bin.AG.class))

frag_regions  %>% group_by(Region) %>%
summarize(100-min(bin.AG.class))

frag_regions  %>% group_by(Region) %>%
summarize(100-max(bin.AG.class))


```

```{r AppendixS6 Correlations Between Predictor Variables}
predictor_info <- read.csv(file="Final_Regions_and_Predictors.csv", header=TRUE, sep=",")

# Overall
cor(predictor_info[, c("EVI_mean","bin.AG.class")])

# Biogeographic Region Grouped
cluster1 <- subset(predictor_info,Region=="Lake Turkana")
cluster2 <- subset(predictor_info,Region=="Southern Drylands")
cluster3 <- subset(predictor_info,Region=="Victoria Basin & Albertine Rift Forests")
cluster4 <- subset(predictor_info,Region=="Sudanian Savanna")
cluster5 <- subset(predictor_info,Region=="Subequatorial Savannas & Woodlands")
cluster6 <- subset(predictor_info,Region=="Congolian Forests")

# Regions
cor(cluster1[, c("EVI_mean","bin.AG.class")])
cor(cluster2[, c("EVI_mean","bin.AG.class")])
cor(cluster3[, c("EVI_mean","bin.AG.class")])
cor(cluster4[, c("EVI_mean","bin.AG.class")])
cor(cluster5[, c("EVI_mean","bin.AG.class")])
cor(cluster6[, c("EVI_mean","bin.AG.class")])

```
